A Study of Common Bug Fix Patterns in Rust \thanks{Competing Interests: We have no competing interests and are funded by a Discovery Grant from Canada's Natural Science and Engineering Research Council.}

### 2023-11-12 response.tex

We also added a paragraph on bug classification at the beginning of Section 5.1. It adds three new references. The first is a 1975 study conducted by Enders et al. \cite{endres1975analysis}, which examines common error patterns in the DOS/VS operating system (written in the DOS Macro Assembler langage). The second study is by Knuth~\cite{knuth89:_error_tex}, surveying errors that he added in the development of \TeX. Finally, we added a 1998 study by Flanagan et al. \cite{flanagan1998new}, where they introduce a static debugger for Scheme (a Lisp dialect), centered around the error-prone operators in the language. Additionally, we moved the initial citation of Pan et al~\cite{pan2009toward} up to this paragraph; it studies bugs in seven large-scale Java programs.

\subsection{More discussion on the impact of the evolving Rust specification}

\textit{\textbf{R1:} Additionally, I would appreciate it if the authors could discuss the potential impact of the evolving Rust specification. Given that Rust is a novel programming language, it is subject to rapid grammatical changes, which in turn could influence bug characteristics and symptoms.}

\vspace*{1em} \noindent \textbf{Response:} Yes, the 2022 Rust roadmap\footnote{\url{https://blog.rust-lang.org/inside-rust/2022/04/04/lang-roadmap-2024.html}} includes future proposed changes to Rust. It is indeed a threat to external validity, and we have now discussed it in our threats to external validity (Section 4.3); this threat could be mitigated by refreshing the analysis results once new versions of Rust are broadly used by extant packages.

@@

\textit{Thanks a lot for the information and clarification in the rebuttal and the new paper version. I would suggest the authors make the following changes in the next paper version.}

\subsection{More discussion on the size and type of the patches}

\textit{\textbf{R2:} Inspired by the last paragraph of Section 4.1, can authors discuss which of the following is more important for Rust in the next paper version: fixing bugs and fixing compilation errors?}

\vspace*{1em} \noindent \textbf{Response:} We added a new paragraph at the end of Section 4.1: we consider bugs as potentially more critical. This is because a bug, if left unresolved, can remain dormant within the system and then inadvertently get deployed, whereas a program with compilation errors cannot be shipped by developers.

@@

\textit{\textbf{R2:} In Section 2.3.2, the authors use sentences from an existing paper directly (although quotes are added). I don't think this is a good way to write an academic paper. Please paraphrase these sentences.}

\vspace*{1em} \noindent \textbf{Response:} Paraphrased the three definitions that were directly quoted.

\vspace*{1em} \noindent \textit{\textbf{R2: } I still do not understand Figure 1. I suggest the authors label ASTDiff nodes with code to ease the explanation.}

\vspace*{1em} \noindent \textbf{Response:} We have added the labels {\tt \&tail} and {\tt head.clone()} at the relevant ASTDiff nodes on the tree on the left. The nonterminals themselves don't directly correspond to code---it would be the terminals which are not present in the paths. We have also added more text to the caption of Figure 1 and point at the ExprReference in Path 1 in the text just below the Figure.

@@

\subsection{Minor issues}

\textit{\textbf{R2:} Line 21 on page 8: work around $\rightarrow$ workaround}

@@

\vspace*{1em} \noindent \textbf{Response:} Fixed.

\vspace*{1em} \noindent \textit{\textbf{R2:} Line 49 on page 8: which path is colored in purple in figure 1?}

\vspace*{1em} \noindent \textbf{Response:} It is the rightmost path. We changed the color to blue so it is clearer and added a reference to ``Path 3'' to the text, to match with the Figure.

### 2023-11-12 discussion.tex
reread the edits

While comparing the importance of bugs and compilation errors can be challenging, we tend to consider bugs as potentially more critical. This is because a bug, if left unresolved, can remain dormant within the system and inadvertently get deployed, whereas a program with compilation errors cannot be shipped by developers.

>>>>

To compare the importance of bugs versus compilation errors: we consider bugs as potentially more critical. This is because a bug, if left unresolved, can remain dormant within the system and then inadvertently get deployed, whereas a program with compilation errors cannot be shipped by developers.

### 2023-11-12 methodology.tex
reread the edits

    \caption{\label{fig:extraction}The ASTDiff (left) corresponds to the code change in the yellow box (top right). A DFS of the ASTDiff finds three paths of nonterminals (Paths 1--3), shown in three different colors at right. The first path summarizes the addition of the ExprReference {\tt \&} before tail, while the second and third paths indicate the change from a read of \texttt{head} to a method call, and a call to \texttt{clone()}, respectively.}

@@

In our example, we add a \verb+&+ borrow operator to the variable \verb+tail+, which gives rise to an ASTDiff path introducing a new parent node (\verb+ExprReference+) for \verb+tail+. We also add a call to \verb+clone()+. This addition gives rise to two ASTDiff paths: one showing that the \verb+head+ node now belongs to an \verb+ExprMethodCall+, and one indicating the new sibling node for \verb+head+ representing the \verb+clone()+. (Among the terminals, we only keep \verb+clone+ in our paths, as it is a BC-related keyword.) One path for the \verb+&+ borrow operator and two for the call to \verb+clone()+ result in three paths in the ASTDiff, even if it only looks like we had made two modifications.

@@

\begin{itemize}
    \item bug-fix: The code change rectifies the undesired behavior in the software system, providing fixes for a specific bug type. We also include here changes that improve program performance, as they fix the behaviour of the software.

>>>>

    \caption{\label{fig:extraction}The ASTDiff (left) corresponds to the code change in the yellow box (top right). A DFS of the ASTDiff finds three paths of nonterminals (Paths 1--3), shown in three different colors at right. The paths summarize (1) the addition of the ExprReference {\tt \&} before tail; (2) the change from a read of \texttt{head} to a method call (represented by ExprMethodCall), and (3) specifically a call to distinguished method \texttt{clone()}.}

@@

In our example, we add a \verb+&+ borrow operator to the variable \verb+tail+, which gives rise to an ASTDiff path introducing a new parent node (\verb+ExprReference+) for \verb+tail+. We also add a call to \verb+clone()+. This addition gives rise to two ASTDiff paths: one showing that the \verb+head+ node now belongs to an \verb+ExprMethodCall+, and one indicating the new sibling node for \verb+head+ representing the \verb+clone()+. (Among the terminals, we only keep the distinguished terminal \verb+clone+ in our paths, as it is a BC-related keyword.) One path for the \verb+&+ borrow operator and two for the call to \verb+clone()+ result in three paths in the ASTDiff, even if it only looks like we had made two modifications.

@@

\begin{itemize}
    \item bug-fix: The code change rectifies undesired behavior in the software system, providing fixes for a specific bug type. We also include here changes that improve program performance, as they fix the behaviour of the software.

### 2023-11-12 related.tex
reread the edits

Automated program repair (APR) aims to debug faulty source code without human involvement, sometimes using test cases to guide the repair. Before a repair tool modifies code, it will typically use a fault localization module to find fault locations. A fault localization module ranks program statements based on suspiciousness; one example is Tarantula by \cite{jones2005empirical}. Suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\citep{naish2009spectral,xie2013theoretical}.

\subsection{Automated Program Repair}
We next discuss a number of approaches for automated program repair. First, however, we discuss previous work that aimed to classify bugs---program repair must have a model of what it is attempting to repair in the first place. We then present two mindsets for program repair: search-based and pattern-based. Finally, we survey related work in automated patch correctness assessment, which aims to ensure that the generated changes are valid.

\paragraph{Bug Classification.}
A foundational bug study dates back to~\cite{endres1975analysis}, which classified the bugs that were fixed in a particular release of IBM's DOS/VS operating system, written in the DOS Macro Assembler language. Endres contended that categorizing bugs can aid in uncovering error causes, fixing them, and preventing their recurrence. The Group B of bugs in that work include, among others, bugs that could be prevented by the use of a different programming language, e.g. ``assignment, loading, or saving of address registers forgotten,'' which is somewhat similar to our Rust-specific bug patterns. Later on, \cite{knuth89:_error_tex} analyzed the errors that he made in the development of \TeX, written in the WEB literate programming language. This was perhaps one of the first works in this domain where there is a publicly-available artifact (the \TeX~source code) supporting the data (i.e. the bugs identified), unlike the proprietary DOS/VS code. Knuth includes 15 categories of errors; the most relevant ones to this work are ``language liabilities,'' ``mismatches between modules,'' ``reinforcements of robustness,'' and ``surprising scenarios.'' \cite{flanagan1998new} described MrSpidey, an interactive static debugger for Scheme (a dialect of Lisp). MrSpidey is a front-end to a ESC-style static verification engine that aims to detect bug patterns where operators are passed invalid arguments. The engine relies on program invariants and presents information about attempted proofs of the invariants to the developer. Most recently, \cite{pan2009toward} analyzed seven large-scale widely used Java projects and manually extracted 27 common bug fix patterns.

@@

As previously cited, \cite{pan2009toward}, used a bug fix pattern extractor tool to automatically parse and detect bug patterns within the bug fix hunks with less than seven statements. They ignored larger bug fix hunks as such hunks tended to exhibit random changes and did not have observable meaningful patterns. Though they did the analysis on Java projects, their reported patterns were not specific to Java. \cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method. In their empirical evaluation, they concluded that choosing repair actions probabilistically (weighted by frequency) can help with reducing the search space, hence creating a more effective repair tool.

>>>>

Automated program repair (APR) aims to debug faulty source code without human involvement, sometimes using test cases to guide the repair. Before a repair tool modifies code, it will typically use a fault localization module to find fault locations. The fault localization modules that we are aware of rank program statements based on suspiciousness; one example is Tarantula by \cite{jones2005empirical}. Often, suspiciousness is calculated using a statistical model which relies on the observation that buggy statements are executed mostly by failed test cases~\citep{naish2009spectral,xie2013theoretical}.

\subsection{Automated Program Repair}
We next discuss a number of approaches for automated program repair. First, however, we discuss previous work that aimed to classify bugs---program repair must have a model of what it is attempting to repair in the first place. We then present two mindsets for program repair: search-based and pattern-based. Finally, we survey related work in automated patch correctness assessment, which aims to ensure that changes by automated repair tools are valid.

\paragraph{Bug Classification.}
A foundational bug study dates back to~\cite{endres1975analysis}, which classified the bugs that were fixed in a particular release of IBM's DOS/VS operating system, written in the DOS Macro Assembler language. Endres contended that categorizing bugs can aid in uncovering error causes, fixing them, and preventing their recurrence. The Group B of bugs in that work include, among others, bugs that could be prevented by the use of a different programming language, e.g. ``assignment, loading, or saving of address registers forgotten,'' which is somewhat similar to our Rust-specific bug patterns. Later on, \cite{knuth89:_error_tex} analyzed the errors that he made in the development of \TeX, written in the WEB literate programming language. This was perhaps one of the first works in this domain where there is a publicly-available artifact (the \TeX~source code) supporting the data (i.e. the bugs identified), unlike the proprietary DOS/VS code. Knuth includes 15 categories of errors; the most relevant ones to this work are ``language liabilities,'' ``mismatches between modules,'' ``reinforcements of robustness,'' and ``surprising scenarios.'' \cite{flanagan1998new} described MrSpidey, an interactive static debugger for Scheme (a dialect of Lisp). MrSpidey is a front-end to an ESC-style static verification engine which aims to detect bug patterns where operators are passed invalid arguments. The engine relies on program invariants and presents information about attempted proofs of the invariants to the developer. Most recently, \cite{pan2009toward} analyzed seven large-scale widely used Java projects and manually extracted 27 common bug fix patterns.

@@

\cite{pan2009toward} used a bug fix pattern extractor tool to automatically parse and detect bug patterns within bug fix hunks with fewer than seven statements. They ignored larger bug fix hunks---such hunks tended to exhibit random changes rather than having patterns that they could derive meaning from. Though they analyzed Java projects, their reported patterns were not specific to Java. \cite{martinez2015mining,martinez2012mining} exploited the frequency of observed patterns to introduce a heuristic patch searching method. In their empirical evaluation, they concluded that choosing repair actions probabilistically (weighted by frequency) can help with reducing the search space, hence creating a more effective repair tool.
